---
title: "Mali_Data_Analysis"
author: "Winnie Kulei"
date: "Date ~ `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true 
    toc_depth: 6
    number_sections: true
    theme: united
    highlight: tango
    
---

<style type="text/css">

h1.title {
  font-size: 20px;
  color: Blue;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 16px;
  font-family: "Source Sans Pro Semibold", Times, serif;
  color: Blue;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 16px;
  font-family: "Source Sans Pro Semibold", Times, serif;
  color: Blue;
  text-align: center;
}
</style>



</br>

</br>


***
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,error = T)

#Clear the memory
rm(list  =  ls())

#Import packages
library(tidyverse)
library(readxl)
library(magrittr)
library(broom)
library(quickr)
library(wordstonumbers)
```

# List Selected Sectors in Mali

```{r}

sectors <- c("Bougouni", "Koutiala", "Sikasso")


```
# Single Phase Product {.tabset}

## Read Excel files containing data

```{r}

monthlyrf_files <- list.files(path = "C:\\Users\\Winnie Kulei\\Documents\\DAVU AI\\Mali_JJA_Monthly_summary", pattern = "*.xls", full.names = T)
mali_monthly_data <- setNames(sapply(monthlyrf_files, read_excel, simplify = F), sectors) %>%
  bind_rows(.id =  "Stations")
glimpse(mali_monthly_data)

```
### Calculate total in-season rainfall

```{r}

mali_monthly_data <- mali_monthly_data %>%
  mutate(Total_rain = select(., Jun:Sep) %>% apply(1, sum, na.rm = T))


#Drop the months columns
mali_monthly_data <- mali_monthly_data %>% select(-Jun, -Jul, -Aug, -Sep)


#Rearrange columns
mali_monthly_data %<>%
  select(Stations,Years, Total_rain, Yield) %>%
  arrange(Stations, Years)
glimpse(mali_monthly_data)

```

## Time series plots per sector

```{r}

monthlyrf_plot <- ggplot(mali_monthly_data, aes(x = Years))
monthlyrf_plot <- monthlyrf_plot + geom_line(aes(y = Total_rain, color = "Rainfall")) + geom_line(aes(y = Yield, color = "Yield")) + facet_wrap(facets = vars(Stations))
monthlyrf_plot

```

Drop rows with missing values

```{r}

mali_monthly_data <- mali_monthly_data %>% drop_na()
```

## Functions
```{r}
SatisfactionIndex1 <- function(Payout_diff){
  return((1 - Payout_diff) * 10)
}

RiskPremium <- function(Total_payout,Years){
  return(Total_payout/Years)
}

```
## Sector linear regression and contract evaluation {.tabset}

### Bougouni
```{r}

mali_monthly_data_bougouni <- mali_monthly_data %>% filter(Stations == "Bougouni", between(Years, 1990,2013))
dim(mali_monthly_data_bougouni)

mean_yield_bougouni <- mean(mali_monthly_data_bougouni$Yield)

mali_monthly_data_bougouni$Yield_loss <- ifelse(mali_monthly_data_bougouni$Yield >= mean_yield_bougouni, 0, (mean_yield_bougouni - mali_monthly_data_bougouni$Yield) / mean_yield_bougouni*100)

plot_bougouni <- ggplot(mali_monthly_data_bougouni, aes(x = Total_rain, y = Yield_loss)) +geom_point() + geom_smooth(method = "lm")
plot_bougouni

linear_mod_bougouni <- lm(Yield_loss~Total_rain, data = mali_monthly_data_bougouni)
summary(linear_mod_bougouni)
plot(linear_mod_bougouni)

mali_monthly_data_bougouni <- mali_monthly_data_bougouni[-c(1,2,6),]
dim(mali_monthly_data_bougouni)

linear_mod_bougouni <- lm(Yield_loss~Total_rain, data = mali_monthly_data_bougouni)
plot(linear_mod_bougouni)

mod_res_bougouni1 = tidy(linear_mod_bougouni) %>% 
  mutate(Sig_at_95_CI = ifelse(p.value < 0.05, "Significant","Not significant"),
       Sig_at_90_CI = ifelse(p.value < 0.1, "Significant","Not significant"),
       estimate = round(estimate,2),
         std.error = round(std.error,2),
         statistic = round(statistic,2),
         p.value = round(p.value,2)) %>% 
  arrange(desc(Sig_at_95_CI)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni1,names(mod_res_bougouni1))

mod_res_bougouni2 <- glance(linear_mod_bougouni) %>%
  transmute(r.squared = round(r.squared,2),
            adj.r.squared = round(adj.r.squared,2)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni2,names(mod_res_bougouni2))




# Contract evaluation

intercept_bougouni <- linear_mod_bougouni$coefficients[1]
mod_coef_bougouni <- linear_mod_bougouni$coefficients[2]

rain_at_zero_loss_bougouni <- as.numeric(-intercept_bougouni/mod_coef_bougouni)
rain_at_zero_loss_bougouni

rain_at_max_loss_bougouni <- mali_monthly_data_bougouni$Total_rain[14]
rain_at_max_loss_bougouni
  
max_loss_bougouni <- as.numeric(mod_coef_bougouni*rain_at_max_loss_bougouni + intercept_bougouni)
  
payout_rate_bougouni <- 100/(rain_at_zero_loss_bougouni - rain_at_max_loss_bougouni)
payout_rate_bougouni


contract_bougouni <- mali_monthly_data_bougouni %>%
  mutate(Actual_payout = ifelse(mali_monthly_data_bougouni$Total_rain >= rain_at_zero_loss_bougouni, 0, (rain_at_zero_loss_bougouni -mali_monthly_data_bougouni$Total_rain) * payout_rate_bougouni)) %>%
  mutate(Actual_payout = ifelse(Actual_payout >= 100, 100, Actual_payout)) %>%
  mutate(Payout_diff = abs(Actual_payout - Yield_loss)/ 100) %>%
  mutate(Sat_index = SatisfactionIndex1(Payout_diff))

glimpse(contract_bougouni)

sat_index_bougouni <- mean(contract_bougouni$Sat_index)
sat_index_bougouni

```
$$y = -0.0113x + 12.3907$$

Outliers removed

```{r}

mali_monthly_data_bougouni <- mali_monthly_data_bougouni[-c(1,2,6),]
dim(mali_monthly_data_bougouni2)

mean_yield_bougouni2 <- mean(mali_monthly_data_bougouni2$Yield)
mean_yield_bougouni2
#median(mali_monthly_data_bougouni2$Yield)

mali_monthly_data_bougouni2$Yield_loss <- ifelse(mali_monthly_data_bougouni2$Yield >= mean_yield_bougouni2,0, (mean_yield_bougouni2 - mali_monthly_data_bougouni2$Yield) / mean_yield_bougouni2*100)

linear_mod_bougouni2 <- lm(Yield_loss~Total_rain, data = mali_monthly_data_bougouni2)
#summary(linear_mod_bougouni2)
plot(linear_mod_bougouni2)

mod_res_bougouni1 = tidy(linear_mod_bougouni2) %>% 
  mutate(Sig_at_95_CI = ifelse(p.value < 0.05, "Significant","Not significant"),
       Sig_at_90_CI = ifelse(p.value < 0.1, "Significant","Not significant"),
       estimate = round(estimate,2),
         std.error = round(std.error,2),
         statistic = round(statistic,2),
         p.value = round(p.value,2)) %>% 
  arrange(desc(Sig_at_95_CI)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni1,names(mod_res_bougouni1))

mod_res_bougouni2 <- glance(linear_mod_bougouni2) %>%
  transmute(r.squared = round(r.squared,2),
            adj.r.squared = round(adj.r.squared,2)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni2,names(mod_res_bougouni2))

plot_bougouni2 <- ggplot(mali_monthly_data_bougouni2, aes(x = Total_rain, y = Yield_loss)) +geom_point() + geom_smooth(method = "lm")
plot_bougouni2


# Contract evaluation

intercept_bougouni2 <- linear_mod_bougouni2$coefficients[1]
mod_coef_bougouni2 <- linear_mod_bougouni2$coefficients[2]

rain_at_zero_loss_bougouni2 <- as.numeric(-intercept_bougouni2/mod_coef_bougouni2)
rain_at_zero_loss_bougouni2

rain_at_max_loss_bougouni2 <- mali_monthly_data_bougouni2$Total_rain[14]
  
max_loss_bougouni2 <- as.numeric(mod_coef_bougouni2*rain_at_max_loss_bougouni2 + intercept_bougouni2)
  
payout_rate_bougouni2 <- 100/(rain_at_zero_loss_bougouni2 - rain_at_max_loss_bougouni2)
payout_rate_bougouni2


contract_bougouni2 <- mali_monthly_data_bougouni2 %>%
  mutate(Actual_payout = ifelse(mali_monthly_data_bougouni2$Total_rain >= rain_at_zero_loss_bougouni2, 0, (rain_at_zero_loss_bougouni2 -mali_monthly_data_bougouni2$Total_rain) * payout_rate_bougouni2)) %>%
  mutate(Actual_payout = ifelse(Actual_payout >= 100, 100, Actual_payout)) %>%
  mutate(Payout_diff = abs(Actual_payout - Yield_loss)/100) %>%
  mutate(Sat_index = SatisfactionIndex1(Payout_diff))

glimpse(contract_bougouni2)

sat_index_bougouni2 <- mean(contract_bougouni2$Sat_index)
sat_index_bougouni2

```
$$y = -0.025x + 20.2957$$
Using estimated yield

```{r}

mali_monthly_data_bougouni3 <- mali_monthly_data_bougouni %>%
  mutate(Yield <- predict(loess(mali_monthly_data_bougouni$Yield~mali_monthly_data_bougouni$Total_rain)))

mean_yield_bougouni3 <- mean(mali_monthly_data_bougouni3$Yield)

mali_monthly_data_bougouni3$Yield_loss <- ifelse(mali_monthly_data_bougouni3$Yield >= mean_yield_bougouni3, 0, (mean_yield_bougouni3 - mali_monthly_data_bougouni3$Yield) / mean_yield_bougouni3*100)

linear_mod_bougouni3 <- lm(Yield_loss~Total_rain, data = mali_monthly_data_bougouni3)
#summary(linear_mod_bougouni3)
plot(linear_mod_bougouni3)

mod_res_bougouni1 = tidy(linear_mod_bougouni3) %>% 
  mutate(Sig_at_95_CI = ifelse(p.value < 0.05, "Significant","Not significant"),
       Sig_at_90_CI = ifelse(p.value < 0.1, "Significant","Not significant"),
       estimate = round(estimate,2),
         std.error = round(std.error,2),
         statistic = round(statistic,2),
         p.value = round(p.value,2)) %>% 
  arrange(desc(Sig_at_95_CI)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni1,names(mod_res_bougouni1))

mod_res_bougouni2 <- glance(linear_mod_bougouni3) %>%
  transmute(r.squared = round(r.squared,2),
            adj.r.squared = round(adj.r.squared,2)) %>% as.data.frame()

quickr::decorate_table(mod_res_bougouni2,names(mod_res_bougouni2))

plot_bougouni3 <- ggplot(mali_monthly_data_bougouni3, aes(x = Total_rain, y = Yield_loss)) +geom_point() + geom_smooth(method = "lm")
plot_bougouni3


# Contract evaluation

intercept_bougouni3 <- linear_mod_bougouni3$coefficients[1]
mod_coef_bougouni3 <- linear_mod_bougouni3$coefficients[2]

rain_at_zero_loss_bougouni3 <- as.numeric(-intercept_bougouni3/mod_coef_bougouni3)
rain_at_zero_loss_bougouni3

rain_at_max_loss_bougouni3 <- mali_monthly_data_bougouni3$Total_rain[2]
rain_at_max_loss_bougouni3
  
max_loss_bougouni3 <- as.numeric(mod_coef_bougouni3*rain_at_max_loss_bougouni3 + intercept_bougouni3)
  
payout_rate_bougouni3 <- 100/(rain_at_zero_loss_bougouni3 - rain_at_max_loss_bougouni3)
payout_rate_bougouni3


contract_bougouni3 <- mali_monthly_data_bougouni3 %>%
  mutate(Actual_payout = ifelse(mali_monthly_data_bougouni3$Total_rain >= rain_at_zero_loss_bougouni3, 0, (rain_at_zero_loss_bougouni3 -mali_monthly_data_bougouni3$Total_rain) * payout_rate_bougouni3)) %>%
  mutate(Actual_payout = ifelse(Actual_payout >= 100, 100, Actual_payout)) %>%
  mutate(Payout_diff = abs(Actual_payout - Yield_loss)/ 100) %>%
  mutate(Sat_index = SatisfactionIndex1(Payout_diff))

glimpse(contract_bougouni3)

sat_index_bougouni3 <- mean(contract_bougouni3$Sat_index)
sat_index_bougouni3

```

Premium calculation
```{r}

Total_payout_bougouni <- sum(contract_bougouni$Actual_payout)
Years_bougouni<- 23

RiskPremium(Total_payout_bougouni, Years_bougouni)
```


### Koutiala
```{r}

mali_monthly_data_koutiala <- mali_monthly_data %>% filter(Stations == "Koutiala", between(Years, 1990,2013))
dim(mali_monthly_data_koutiala)

mean_yield_koutiala <- mean(mali_monthly_data_koutiala$Yield)

mali_monthly_data_koutiala$Yield_loss <- ifelse(mali_monthly_data_koutiala$Yield >= mean_yield_koutiala, 0, (mean_yield_koutiala - mali_monthly_data_koutiala$Yield) / mean_yield_koutiala*100)

linear_mod_koutiala <- lm(Yield_loss~Total_rain, data = mali_monthly_data_koutiala)
summary(linear_mod_koutiala)
plot(linear_mod_koutiala)

mali_monthly_data_koutiala <- mali_monthly_data_koutiala[-c(2,11,15),]
dim(mali_monthly_data_koutiala)

linear_mod_koutiala <- lm(Yield_loss~Total_rain, data = mali_monthly_data_koutiala)
plot(linear_mod_koutiala)

plot_koutiala <- ggplot(mali_monthly_data_koutiala, aes(x = Total_rain, y = Yield_loss)) +geom_point() + geom_smooth(method = "lm")
plot_koutiala

mod_res_koutiala1 = tidy(linear_mod_koutiala) %>% 
  mutate(Sig_at_95_CI = ifelse(p.value < 0.05, "Significant","Not significant"),
       Sig_at_90_CI = ifelse(p.value < 0.1, "Significant","Not significant"),
       estimate = round(estimate,2),
         std.error = round(std.error,2),
         statistic = round(statistic,2),
         p.value = round(p.value,2)) %>% 
  arrange(desc(Sig_at_95_CI)) %>% as.data.frame()

quickr::decorate_table(mod_res_koutiala1,names(mod_res_koutiala1))

mod_res_koutiala2 <- glance(linear_mod_koutiala) %>%
  transmute(r.squared = round(r.squared,2),
            adj.r.squared = round(adj.r.squared,2)) %>% as.data.frame()

quickr::decorate_table(mod_res_koutiala2,names(mod_res_koutiala2))


```

### Sikasso
```{r}


mali_monthly_data_sikasso <- mali_monthly_data %>% filter(Stations == "Sikasso", between(Years, 1990,2013))
dim(mali_monthly_data_sikasso)

mean_yield_sikasso <- mean(mali_monthly_data_sikasso$Yield)

mali_monthly_data_sikasso$Yield_loss <- ifelse(mali_monthly_data_sikasso$Yield >= mean_yield_sikasso, 0, (mean_yield_sikasso - mali_monthly_data_sikasso$Yield) / mean_yield_sikasso*100)

linear_mod_sikasso <- lm(Yield_loss~Total_rain, data = mali_monthly_data_sikasso)
summary(linear_mod_sikasso)
plot(linear_mod_sikasso)

mali_monthly_data_sikasso <- mali_monthly_data_sikasso[-c(1,4,5),]
dim(mali_monthly_data_sikasso)

linear_mod_sikasso <- lm(Yield_loss~Total_rain, data = mali_monthly_data_sikasso)
plot(linear_mod_sikasso)

plot_sikasso <- ggplot(mali_monthly_data_sikasso, aes(x = Total_rain, y = Yield_loss)) +geom_point() + geom_smooth(method = "lm")
plot_sikasso

mod_res_sikasso1 = tidy(linear_mod_sikasso) %>% 
  mutate(Sig_at_95_CI = ifelse(p.value < 0.05, "Significant","Not significant"),
       Sig_at_90_CI = ifelse(p.value < 0.1, "Significant","Not significant"),
       estimate = round(estimate,2),
         std.error = round(std.error,2),
         statistic = round(statistic,2),
         p.value = round(p.value,2)) %>% 
  arrange(desc(Sig_at_95_CI)) %>% as.data.frame()

quickr::decorate_table(mod_res_sikasso1,names(mod_res_sikasso1))

mod_res_sikasso2 <- glance(linear_mod_sikasso) %>%
  transmute(r.squared = round(r.squared,2),
            adj.r.squared = round(adj.r.squared,2)) %>% as.data.frame()

quickr::decorate_table(mod_res_sikasso2,names(mod_res_sikasso2))



```

# Multi-trigger Product

## Read Excel files containing data
```{r}

mali_rf <-  read_xlsx("C:\\Users\\Ebimoboere Claudius\\Documents\\ICRISAT-MANOBI\\PLUIES.xlsx")
#glimpse(mali_rf)

#mali_rf <- mali_rf %>% gather(Day, Rain, One:ThirtyOne)

phase1_start
phase1_end


phase2_start
phase2_end


phase3_start
phase3_end


ds_start
ds_end


er_start
er_end

#Monthly and decadal rainfall distribution plots for each sector

```


## Extract JJAS data for Bougouni
```{r}
#Bougouni StationID - 27016200

mali_rf_bougouni <- mali_rf %>% filter(StationID == "27016200", between(Years, 1990,2013), between(Month, 6,9))
dim(mali_rf_bougouni)

mali_rf_bougouni2 <- mali_rf_bougouni %>% gather(Day, Rain, One:ThirtyOne)

mali_rf_bougouni3 <- mali_rf_bougouni2 %>% arrange(Years,Month)
#mali_rf_bougouni3 <- mali_rf_bougouni %>% mali_rf_bougouni$Day <- 1:31
glimpse(mali_rf_bougouni3)


#words_to_numbers("three billion")



mali_rf_bougouni2 <- mali_rf_bougouni %>%
  group_by(StationID,Years, Month) %>%
  mutate(pivot_longer(-Month, cols = One:ThirtyOne, names_to = "Day", values_to = "Rain"))

mali_rf_bougouni2 <- mali_rf_bougouni %>%
  pivot_longer(mali_rf_bougouni, cols = One:ThirtyOne, names_to = "Day", values_to = "Rain")


monthlyrf_bougouni <- mali_rf_bougouni %>%
  mutate(Monthly_rain = select(., One:ThirtyOne) %>% apply(1, sum, na.rm = T))
monthlyrf_bougouni <- monthlyrf_bougouni %>%
  select(-One:-ThirtyOne) %>%
  group_by(Years) %>% summarise(JJAS_rain = sum(Monthly_rain))

#yield_bougouni <- mali_monthly_data_bougouni %>%
 # select(Yield)

monthlyrf_bougouni <- monthlyrf_bougouni %>%
  mutate(Yield = mali_monthly_data_bougouni$Yield)

#monthlyrf_bougouni$Yield <- yield_bougouni

#single_phase_bougouni <- full_join(monthlyrf_bougouni,yield_bougouni, by = character())

```

## Extract JJAS data for Koutiala
```{r}
#Koutiala StationID - 27014400

mali_rf_koutiala <- mali_rf %>% filter(StationID == "27014400", between(Years, 1990,2013), between(Month, 6,9))
dim(mali_rf_koutiala)


monthlyrf_koutiala <- mali_rf_koutiala %>%
  mutate(Monthly_rain = select(., One:ThirtyOne) %>% apply(1, sum, na.rm = T))
monthlyrf_koutiala <- monthlyrf_koutiala %>%
  select(-One:-ThirtyOne) %>%
  group_by(Years) %>% summarise(JJAS_rain = sum(Monthly_rain))

#yield_koutiala <- mali_monthly_data_koutiala %>%
 # select(Yield)

monthlyrf_koutiala <- monthlyrf_koutiala %>%
  mutate(Yield = mali_monthly_data_koutiala$Yield)

#monthlyrf_koutiala$Yield <- yield_koutiala


```
## Extract JJAS data for Sikasso
```{r}
#Sikasso StationID - 27016500

mali_rf_sikasso <- mali_rf %>% filter(StationID == "27016500", between(Years, 1990,2013), between(Month, 6,9))
dim(mali_rf_sikasso)

monthlyrf_sikasso <- mali_rf_sikasso %>%
  mutate(Monthly_rain = select(., One:ThirtyOne) %>% apply(1, sum, na.rm = T))
monthlyrf_sikasso <- monthlyrf_sikasso %>%
  select(-One:-ThirtyOne) %>%
  group_by(Years) %>% summarise(JJAS_rain = sum(Monthly_rain))

#yield_sikasso <- mali_monthly_data_sikasso %>%
 # select(Yield)

monthlyrf_sikasso <- monthlyrf_sikasso %>%
  mutate(Yield = mali_monthly_data_sikasso$Yield)

#monthlyrf_sikasso$Yield <- yield_sikasso

```

Test
```{r}

mali_rf_sector <- mali_rf %>% filter(StationID == "270167", between(year, 1990,2013), between(Month, 6,9))
dim(mali_rf_sector)
```

